\bfseries Don't interpolate your data!

### 2022-10-23 paper/main.tex

\paragraph{Fake data}
We demonstrate the effectiveness of the spectral combination method on artificial data.
The spectral data were generated from a mean spectral expectation that is a continuum plus a randomly generated (not at all physical) line list, where each line equivalent width was set by a random draw from a power law.
Each spectral epoch was given a velocity (Doppler shift) by putting the epochs equally spaced on a sinusoidal variation with a semi-amplitude of $30\,\kmps$ (to emulate the motion of the spectrograph with respect to the Solar System barycenter).
The data were generated with an assumption of unresolved spectral lines and a Gaussian line-spread function (LSF) by simply evaluating the spectral expectation model at pixel centers, so the LSF is effectively the pixel-convolved LSF.
After sampling the mean spectral expectation model onto an observational pixel grid, we add Gaussian noise with a uniform variance (HOGG NO WE DON'T!) such that the signal-to-noise in the continuum is XXXX per pixel in each epoch spectrum.
At a rate of XXX pixels were randomly and independently marked as ``bad'' and offset by a large offset.

HOGG: MAKE 2 DATA SETS, one poorly sampled, one well sampled.
The individual pixel grids $x_i$ were set to slightly under-sample the pixel-convolved LSF, and they were fixed in the rest frame of the spectrograph.

At each epoch $i$, there is a spectrum $y_i$ with $M_i=XXX$ pixels, a spectrograph rest-frame pixel grid $x_i$ (in log wavelength $x$), and a $M_i$-pixel binary bad-pixel mask $b_i$.
The full data set is shown in Figure~\label{fig:data1}, and a zoom-in on one epoch is shown in \figurename~\ref{fig:data2}.
\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=1.3\figurewidth]{notebooks/data1.pdf}\\
    \includegraphics[width=\figurewidth]{notebooks/datazoom.pdf}
    \end{center}
    \caption{The top 8 panels show the $N=8$ epochs of the poorly sampled input data. The bad pixels are indicated with grey bars. The bottom 2 panels show a zoom in on one of the epochs, along with the true spectrum used to generate the data. The true spectrum is shown at zero Doppler shift, while each epoch spectrum is at a finite Doppler shift $\Delta x_i$.\label{fig:data1}}
    \end{mdframed}
\end{figure}
\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=1.3\figurewidth]{notebooks/data2.pdf}
    \end{center}
    \caption{Similar to the top panels of \figurename~\ref{fig:data1} except showing the well sampled input data.\label{fig:data2}}
    \end{mdframed}
\end{figure}

For the combined spectrum we choose a pixel grid $x_\star$ with $M_\star$ pixels evenly spaced in log wavelength $x$, separated by $1/R$, which (just barely) samples the pixel-convolved LSF well.
We treat every pixel in the data as having the same noise variance $\sigma_y$ (which is indeed how they were generated).

\paragraph{Fiducial case}
HOGG: OUR METHOD NEEDS A CONSISTENT NAME USED THROUGHOUT. FORWARD MODEL??

HOGG: Give full list of input parameters for the Fiducial.

@@

HOGG: HERE SAY WHAT WE FIND RELATIVELY to Standard Practice (tm).

\paragraph{Standard Practice (tm)}
In standard practice, the individual spectra $y_i$ are interpolated to rest-frame spectra $y'_i$ using an interpolator.
In detail, the data are shifted and also resampled onto the output pixel grid $x_\star$.
The interpolator is a choice; cubic spline, Lanczos, and sinc interpolations are common, since they have good properties with respect to the spectrograph band limit.
Here we choose a cubic spline for convenience ANDY: WHAT SHOULD WE CHOOSE??.

The interpolated spectra $y'_i$ are, by construction, all on the same rest-frame wavelength grid $x_\star$ so they can be averaged, with a straight average, a median, an inverse-variance-weighted mean, or any more sophisticated algorithm.
Here we choose an inverse-variance-weighted mean, but zeroing out the inverse variance at the locations of the bad pixels.

@@

@@

>>>>

\paragraph{Fake data}
We demonstrate the effectiveness of the spectral combination method on artificial data.
The spectral data were generated from a mean spectral expectation that is a continuum plus a randomly generated (not at all physical) line list, where each line equivalent width was set by a random draw from a power law.
Each spectral epoch was given a velocity (Doppler shift) by putting the epochs equally spaced on a sinusoidal variation with a semi-amplitude of $30\,\kmps$ (to emulate the motion of the spectrograph with respect to the Solar System barycenter).
The spectral expectation model is built on the assumption of a normalized (unit) continuum and unresolved spectral lines with a Gaussian line-spread function (LSF) with one-sigma width $1/R$, where $R=135\,000$ is the spectrograph resolution).

The sampled data at each epoch were made by simply evaluating the spectral expectation model (shifted by the Doppler shift) at the pixel centers, such that the $R=135\,000$ LSF is effectively the pixel-convolved LSF.
After sampling the mean spectral expectation model onto an observational pixel grid, Gaussian noise is added with a variance that grows linearly with expected flux, such that the signal-to-noise per pixel in the continuum has a definite known value in each epoch spectrum.
At a rate of 0.01 (one percent), sets of three adjacent pixels were randomly and independently marked as ``bad'' and offset by a large positive offset.

According to these rules we made two input data sets.
One is poorly sampled, with $M=171$ spectral pixels in the input data separated by $2/R$ in $x$ (natural logarithm of wavelength), and signal-to-noise per pixel of 17 in the continuum.
The other is well sampled, with $M=340$ spectral pixels in the input data separated by $1/R$ in $x$, and signal-to-noise per pixel of 12 in the continuum.
In each case there are $N=8$ epochs.
The two complete data sets are shown in \figurename~\label{fig:data1} and \figurename~\ref{fig:data2}.
\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=1.3\figurewidth]{notebooks/data1.pdf}\\
    \includegraphics[width=\figurewidth]{notebooks/datazoom.pdf}
    \end{center}
    \caption{The top 8 panels show the $N=8$ epochs of the poorly sampled input data, with $N=8$, $M=171$, and signal-to-noise per pixel of 17 in the continuum. The bad pixels are indicated with grey bars. The bottom 2 panels show a zoom in on one of the epochs, along with the true spectrum used to generate the data. The true spectrum is shown at zero Doppler shift, while each epoch spectrum is at a finite Doppler shift $\Delta x_i$ relative to the spectrograph pixel grid.\label{fig:data1}}
    \end{mdframed}
\end{figure}
\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=1.3\figurewidth]{notebooks/data2.pdf}
    \end{center}
    \caption{Similar to the top panels of \figurename~\ref{fig:data1} except showing the well sampled input data with $N=8$, $M=340$, and signal-to-noise per pixel of 12 in the continuum.\label{fig:data2}}
    \end{mdframed}
\end{figure}

\paragraph{Forward Model (tm)}
HOGG: Give full list of input parameters for the Fiducial. NOTE THAT WE DON'T DO THE NOISE IVAR RIGHT.

@@

\paragraph{Standard Practice (tm)}
In standard practice, the individual spectra $y_i$ are interpolated to rest-frame spectra $y'_i$ using an interpolator.
In detail, the data are shifted and also resampled onto the output pixel grid $x_\star$.
The interpolator is a choice; cubic spline, Lanczos, and sinc interpolations are common, since they have good properties with respect to the spectrograph band limit.
Here we choose a cubic spline for convenience, but there are many other choices, including linear, sinc, and Lanczos interpolators, among many others.

The interpolated spectra $y'_i$ are, by construction, all on the same rest-frame wavelength grid $x_\star$ so they can be averaged, with a straight average, a median, an inverse-variance-weighted mean, or any more sophisticated algorithm.
Here we choose an inverse-variance-weighted mean (HOGG: NO WE DON'T), but zeroing out the inverse variance at the locations of the bad pixels.

@@

HOGG: HERE SAY WHAT WE FIND RELATIVELY between the Forward Model (tm) and Standard Practice (tm).

@@

HOGG: Note that if we choose $P<M$ we start to see noise correlations in the Forward Model (tm) too.
That's one of a few reasons we like $P=M$.
